---
title: "Assignement Exam"
author: "Michele Berti, Gianluca Guglielmo, Niccolò Tosato, Isacco Zinna"
date: "1/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prove

```{r}
#needed for function shift
library(magic)
library(zoo)
```


```{r}
shifted <-function(col, shift=10){
  res = magic::shift(col, i=shift)
  if (shift > 0){
    res[seq.int(from=1, length.out=abs(shift))] = NA
  }
  if (shift < 0){
    res[seq.int(to=length(res), length.out=abs(shift))] = NA # setting the tail of the array to NA
  }
  res
}
```


(just code in this section)
```{r gath, echo = FALSE}
ds <- read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv")



ds.camp <- ds[ds$denominazione_regione == "Campania", ]
rm(ds)
tempd <- data.frame(nuovi_positivi_orig = ds.camp$nuovi_positivi)
ds.camp$data <- as.POSIXct(ds.camp$data, tryFormats = c("%Y-%m-%d %H:%M:%OS",
                          "%Y/%m/%d %H:%M:%OS",
                          "%Y-%m-%d %H:%M",
                          "%Y/%m/%d %H:%M",
                          "%Y-%m-%d",
                          "%Y/%m/%d"))

#cumulativi : deceduti,tamponi,totale_casi,casi_testati
ds.camp$deceduti=c(0,diff(ds.camp$deceduti)) #metto gli attuali al posto del cumulativo
ds.camp$tamponi=c(0,diff(ds.camp$tamponi)) #metto gli attuali al posto del cumulativo
ds.camp$casi_testati=c(0,diff(ds.camp$casi_testati)) #metto gli attuali al posto del cumulativo
colonne=c("data",
          "ricoverati_con_sintomi",#"ricoverati_con_sintomi_variazione",
          "terapia_intensiva",#"terapia_intensiva_variazione",
          #"totale_ospedalizzati",#"totale_ospedalizzati_variazione",
          "isolamento_domiciliare",#"isolamento_domiciliare_variazione",
          #"totale_positivi",#"totale_positivi_variazione",
          "nuovi_positivi",#"nuovi_positivi_variazione",
          "dimessi_guariti",#"dimessi_guariti_variazione",
          "deceduti",#"deceduti_variazione",
          "tamponi",#"tamponi_variazione",
          "casi_testati")#,"casi_testati_variazione")


colonne=colonne[-1]
for(i in colonne) {
  temp=ds.camp[,as.character(i)]
  temp=c(0,diff(temp))
  temp=temp/shifted(ds.camp[,as.character(i)],1)
  temp[is.nan(temp)]=0
  ds.camp[,as.character(paste0(i,"_variazione"))]=temp
}

ds.camp[ds.camp==Inf]=1
ds.camp[ds.camp==-Inf]=-1

#riordino le colonne in maniera accettabile
colonne=c("data",
          "ricoverati_con_sintomi","ricoverati_con_sintomi_variazione",
          "terapia_intensiva","terapia_intensiva_variazione",
          #"totale_ospedalizzati","totale_ospedalizzati_variazione",
          "isolamento_domiciliare","isolamento_domiciliare_variazione",
          #"totale_positivi","totale_positivi_variazione",
          "nuovi_positivi","nuovi_positivi_variazione",
          "dimessi_guariti","dimessi_guariti_variazione",
          "deceduti","deceduti_variazione",
          "tamponi","tamponi_variazione",
          "casi_testati","casi_testati_variazione",
          "walking", "driving")

apple <- read.csv("https://raw.githubusercontent.com/ActiveConclusion/COVID19_mobility/master/apple_reports/apple_mobility_report.csv")
apple.camp <- apple[apple$subregion_and_city == "Campania Region",]

apple.camp <- apple.camp[as.POSIXct(apple.camp$date, format = "%Y-%m-%d") >=  as.POSIXct("2020-02-24", format = "%Y-%m-%d"),]

ds.camp <- ds.camp[ds.camp$data < Sys.Date() - 1,]

ds.camp$walking <- apple.camp$walking
ds.camp$driving <- apple.camp$driving

ds.camp=subset(ds.camp, select = colonne)

colonne=colonne[-1]

for (i in colonne){
  ds.camp[i] <- rollmean(ds.camp[i], k=7, fill=NA, align="center")
}

ds.camp$nuovi_positivi_orig <- 0
tempd <- head(tempd, -2)
ds.camp$nuovi_positivi_orig <- tempd$nuovi_positivi_orig
rm(colonne)
rm(temp)

```


```{r gath, echo = FALSE}



#new range to delete the superfluous day needed for the difference


#ds.camp.range <- ds.camp[ds.camp$data >= 
#                          as.POSIXct("2020-09-01 00:00:00", 
#                          format = "%Y-%m-%d %H:%M:%OS"),]

# Scattered Rt from github
url <- 'https://raw.githubusercontent.com/Biuni/rt-italy/master/src/data/Rt.csv'
rt <- readr::read_csv(url)
rt.camp <- rt[rt$state == "Campania",]
rt.cut <- rt.camp[rt.camp$date > as.POSIXct("2020-08-01", format = "%Y-%m-%d"),]
ds.camp$rt <- 0
length(rt.cut$date)
for (i in 1:length(ds.camp$data)){
  for (j in 1:length(rt.cut$date)){
    if (as.POSIXct(ds.camp$data[i]) < as.POSIXct(rt.cut$date[j])){
      ds.camp$rt[i] <- rt.cut$ML[j]
      break
   }
  }
}

 

# Continuous Rt from Harvard

 

url2 <- "https://dataverse.harvard.edu/api/access/datafile/4346906?format=original&gbrecs=true"
rt3 <- readr::read_csv(url2)
rt3.cut <- rt3[rt3$region == "Campania",]
for (i in 1:length(ds.camp$data)){
  for (j in 1:length(rt3.cut$date)){
    if (as.Date(as.POSIXct(ds.camp$data[i])) == as.Date(as.POSIXct(rt3.cut$date[j]))){
      ds.camp$rt[i] <- rt3.cut$mean[j]
  #    print(j)
  #    print(rt3.cut$mean[j])
  #    print(ds.camp.range$rt[i])
      break
    }
  }
}
rm(rt)
rm(rt.camp)
rm(rt.cut)
rm(rt3)
rm(rt3.cut)
ds.camp$color <- "White"



ds.camp[ds.camp$data >= 
                          as.POSIXct("2020-11-06 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS") & ds.camp$data <=
                          as.POSIXct("2020-11-14 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Yellow"

ds.camp[ds.camp$data >=
                          as.POSIXct("2020-11-15 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS") & ds.camp$data <=
                          as.POSIXct("2020-12-05 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Red"

ds.camp[ds.camp$data >=
                          as.POSIXct("2020-12-06 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS") & ds.camp$data <=
                          as.POSIXct("2020-12-23 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Orange"

ds.camp[ds.camp$data >=   as.POSIXct("2020-12-23 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS") & ds.camp$data <=                              as.POSIXct("2020-12-31 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Red"

ds.camp[ds.camp$data >=   as.POSIXct("2021-01-01 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS") & ds.camp$data <=                              as.POSIXct("2021-01-06 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Orange"

ds.camp[ds.camp$data >=  as.POSIXct("2020-12-25 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS") &
          ds.camp$data <   as.POSIXct("2020-12-26 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Red"

ds.camp[ds.camp$data >
                          as.POSIXct("2021-01-06 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Yellow"

ds.camp[ds.camp$data >=
                          as.POSIXct("2021-01-09 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Yellow"#"Orange"

ds.camp[ds.camp$data >=
                          as.POSIXct("2021-01-11 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Yellow"





ds.camp$color=as.factor(ds.camp$color)
ds.camp$tamp_rate=ds.camp$nuovi_positivi/ds.camp$casi_testati
ds.camp$tamp_rate_variazione=ds.camp$nuovi_positivi_variazione/ds.camp$casi_testati_variazione
```
```{r, echo = TRUE}
library(ltm)
chart.Correlation(ds.camp[, c("nuovi_positivi",
                              "ricoverati_con_sintomi",
                              "isolamento_domiciliare",
                              "terapia_intensiva",
                              "rt",
                              "driving",
                              "tamponi_variazione")])
```

```{r, echo = TRUE}

library(pracma)
ds.camp$difference <- ds.camp$nuovi_positivi_orig - ds.camp$nuovi_positivi

ds.camp$day <- weekdays(as.POSIXct(ds.camp$data))

df3 = data.frame(data = ds.camp$data,
                  day = ds.camp$day,
                  nuovi_positivi_orig = ds.camp$nuovi_positivi_orig)

df3$difference <- 0 
df3$difference = ds.camp$difference
df3$day <- factor(df3$day, levels = c("lunedì", "martedì", "mercoledì", "giovedì",
                                      "venerdì", "sabato", "domenica"))

displacement <- vector(mode = "integer", length = 7)

for (i in 1:length(df3$data)){
  if (!is.na(df3$difference[i])){
  if(strcmp(as.character(df3$day[i]), "lunedì")){
    displacement[1] = displacement[1] + df3$difference[i]
  }
  if(strcmp(as.character(df3$day[i]), "martedì")){
    displacement[2] = displacement[2] + df3$difference[i]
  }
  if(strcmp(as.character(df3$day[i]), "mercoledì")){
    displacement[3] = displacement[3] + df3$difference[i]
  }
  if(strcmp(as.character(df3$day[i]), "giovedì")){
    displacement[4] = displacement[4] + df3$difference[i]
  }
  if(strcmp(as.character(df3$day[i]), "venerdì")){
    displacement[5] = displacement[5] + df3$difference[i]
  }
  if(strcmp(as.character(df3$day[i]), "sabato")){
    displacement[6] = displacement[6] + df3$difference[i]
  }
  if(strcmp(as.character(df3$day[i]),"domenica")){
    displacement[7] = displacement[7] + df3$difference[i]
  }
  count = count + 1
  }
}
displacement <- displacement * 7/count

library(plyr)

library(RColorBrewer)
library(ggplot2)
display.brewer.all()
ggplot(data = df3) +
  geom_point(aes(x = data, y = difference, color = day)) +
  xlim(as.POSIXct("2020-09-01 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
       as.POSIXct("2021-02-14 00:00:00", format = "%Y-%m-%d %H:%M:%OS")) +
       scale_color_manual(values = brewer.pal(7,"Spectral")) +
  ylab("displacement")

```

```{r, echo = TRUE}
#ADD COLORS
range <- ds.camp[ds.camp$data > as.POSIXct("2020-09-01 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]
df2 <- data.frame(data = range$data,
                  nuovi_positivi = range$nuovi_positivi,
                  nuovi_positivi_orig = range$nuovi_positivi_orig)
mdf <- reshape2::melt(df2, id.var = "data")

ggplot(data=mdf) +
  geom_line(aes(x=data, y=nuovi_positivi, color = "Original"), color = "black") +
  geom_line(aes(x=data, y=nuovi_positivi_orig, color = "7-day avg"), color = "red") +
  scale_color_discrete(labels = c("Original" = "black",
                                  "7-day avg" = "red")) +
  xlab("Date") +
  ylab("Nuovi Positivi") +
  xlim(as.POSIXct("2020-09-01 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
       as.POSIXct("2021-02-14 00:00:00", format = "%Y-%m-%d %H:%M:%OS")) 
mdf
ggplot(mdf, aes(x = data, y = value, colour = variable)) + 
    geom_line() +
    scale_color_manual(values=c("black", "red"))
  
    



  #+ geom_text(aes(x = as.POSIXct("2021-01-01 00:00:00", format = "%Y-%m-%d %H:%M:%OS"), y = 2000), label = "Terapia Intensiva", color = "blue", fontface = 0, size = 3, family = "Sans Serif") 

```

```{r, echo = TRUE}
#ADD COLORS
library(ggplot2)

google=read.csv("./2020_IT_Region_Mobility_Report.csv")
google=google[google$sub_region_1=="Campania",]
google$date=as.Date(google$date)
google=google[google$date>="2020-09-1",]
google=google[google$sub_region_2=="",]



google$retail_and_recreation_percent_change_from_baseline <- rollmean(google$retail_and_recreation_percent_change_from_baseline, k=7, fill=NA, align="center")


range <- ds.camp[ds.camp$data > as.POSIXct("2020-09-01 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]
range[174] <- NA
range <- range[-174, ]

range$google <- google$retail_and_recreation_percent_change_from_baseline
ggplot(data=range, aes(x=data, y=google, group=1)) +
  #geom_line(data=range, aes(x=data, y=terapia_intensiva, group=1))+
    geom_rect(aes(xmin=as.POSIXct("2020-09-01 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              xmax=as.POSIXct("2020-11-06 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              ymin=-Inf, ymax=Inf),
              alpha = 0.01,
              fill = "white") +
  geom_rect(aes(xmin=as.POSIXct("2020-11-06 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              xmax=as.POSIXct("2020-11-15 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              ymin=-Inf, ymax=Inf),
              alpha = 0.01,
              fill = "#FFFF99") +
  geom_rect(aes(xmin=as.POSIXct("2020-11-15 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              xmax=as.POSIXct("2020-12-05 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              ymin=-Inf, ymax=Inf),
              alpha = 0.01,
              fill = "red") +
  geom_rect(aes(xmin=as.POSIXct("2020-12-05 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              xmax=as.POSIXct("2020-12-23 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              ymin=-Inf, ymax=Inf),
              alpha = 0.01,
              fill = "orange") +
  geom_rect(aes(xmin=as.POSIXct("2020-12-23 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              xmax=as.POSIXct("2020-12-27 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              ymin=-Inf, ymax=Inf),
              alpha = 0.01,
              fill = "red") +
  geom_rect(aes(xmin=as.POSIXct("2020-12-27 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              xmax=as.POSIXct("2020-12-29 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              ymin=-Inf, ymax=Inf),
              alpha = 0.01,
              fill = "orange") +
  geom_rect(aes(xmin=as.POSIXct("2020-12-29 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              xmax=as.POSIXct("2021-01-03 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              ymin=-Inf, ymax=Inf),
              alpha = 0.01,
              fill = "red") +
  geom_rect(aes(xmin=as.POSIXct("2021-01-03 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              xmax=as.POSIXct("2021-01-04 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              ymin=-Inf, ymax=Inf),
              alpha = 0.01,
              fill = "orange") +
  geom_rect(aes(xmin=as.POSIXct("2021-01-04 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              xmax=as.POSIXct("2021-01-06 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              ymin=-Inf, ymax=Inf),
              alpha = 0.01,
              fill = "red") +
  geom_rect(aes(xmin=as.POSIXct("2021-01-06 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              xmax=as.POSIXct("2021-01-09 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              ymin=-Inf, ymax=Inf),
              alpha = 0.01,
              fill = "#FFFF99") +
  geom_rect(aes(xmin=as.POSIXct("2021-01-09 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              xmax=as.POSIXct("2021-01-11 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              ymin=-Inf, ymax=Inf),
              alpha = 0.01,
              fill = "orange") +
  geom_rect(aes(xmin=as.POSIXct("2021-01-11 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              xmax=as.POSIXct("2021-02-14 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
              ymin=-Inf, ymax=Inf),
              alpha = 0.01,
              fill = "#FFFF99") +
  geom_line() +
  #scale_x_continuous(breaks=1:7,labels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) +
  xlab("Date") +
  ylab("Driving") +
  xlim(as.POSIXct("2020-09-01 00:00:00", format = "%Y-%m-%d %H:%M:%OS"),
       as.POSIXct("2021-02-14 00:00:00", format = "%Y-%m-%d %H:%M:%OS")) +
  geom_hline(yintercept = 0)
  


```




#New color
Trying to assign a new color to the Christmas period to account for people's behaviour

```{r, echo = TRUE}
#shiftare rt

library(DataCombine)

date <- function(dd=1, mm, yy=2020){
  text = paste(yy, "-", mm, "-", dd, " 00:00:00", sep="")
  as.POSIXct(text, format = "%Y-%m-%d %H:%M:%OS")
}



d=ds.camp
d_test = ds.camp
#d_test$color = shifted(d_test$color, 14)
#d_test$rt = shifted(d_test$rt, 14)
d_test$temp = shifted(d_test$ricoverati_con_sintomi, -14)
d_test = d_test[d_test$data <= date(01,02,2021) & d_test$data >= date(1,9,2020), ]

#d$color=shifted(d$color,14)
#d$rt=shifted(d$rt, 14)
d$temp=shifted(d$ricoverati_con_sintomi, -14)
d=d[d$data<=date(26,01,2021) & d$data>=date(1,9,2020), ]

d$rt2 <- 1/d$rt
d_test$rt2 <- 1/d_test$rt


d$old_rt=shifted(d$rt, 14)
d$old_color = shifted(d$color, 14)
d_test$old_rt=shifted(d_test$rt, 14)
d_test$old_color = shifted(d_test$color, 14)
d$old_driving = shifted(d$driving, 14)
d_test$old_driving = shifted(d_test$driving, 14)


m_nuovi_positivi=glm(data=d,
                  formula = nuovi_positivi ~
                  I(terapia_intensiva):old_color +
                  deceduti +
                  old_rt:old_color +
                  I(rt^2)+
                  tamp_rate:old_color +
                  I(tamp_rate^2), 
                  family = Gamma(link="log"))

summary(m_nuovi_positivi)
#d_test$temp2 = shifted(d_test$isolamento_domiciliare, -14)
plot(d_test$nuovi_positivi)
lines(predict.glm(m_nuovi_positivi, d_test, type = "response"), col="blue")
m_nuovi_positivi$temp = shifted(m_nuovi_positivi$fitted.values, 14)
lines(m_nuovi_positivi$temp, col="red")

#1515

#20-6 Purple 1520
#20-6 Red 1539
#20-6 Orange 1543
#20-6 Yellow 1558
#20-6 White 1576
#23-31 Purple 01-06 Orange 1508
#23-31 Red 01-06 Orange 1515
#23-31 Red 01-06 Orange but important days Ora 1515
#23-31 Red 01-06 Purple 1520
#23-31 Orange 01-06 Red 1543
#20-6 Orange but important days red 1666
#20-6 Orange but important days orange 1525
#20-6 Orange but important days yellow 1525
#20-6 Red but important days yellow 1537
#20-6 Red but important days orange 1536
#actual dates 1537


m_nuovi_positivi=glm(data=d,
                  formula = nuovi_positivi ~
                  I(terapia_intensiva):old_color +
                  deceduti +
                  old_rt:old_color +
                  old_driving +
                  I(rt^2)+
                  tamp_rate:old_color +
                  I(tamp_rate^2), 
                  family = Gamma(link="log"))

summary(m_nuovi_positivi)
#d_test$temp2 = shifted(d_test$isolamento_domiciliare, -14)
plot(d_test$nuovi_positivi)
lines(predict.glm(m_nuovi_positivi, d_test, type = "response"), col="blue")
m_nuovi_positivi$temp = shifted(m_nuovi_positivi$fitted.values, 14)
lines(m_nuovi_positivi$temp, col="red")

#1516

```








ds.camp[ds.camp$data >=
                          as.POSIXct("2020-12-24 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS") & ds.camp$data <=
                          as.POSIXct("2020-12-27 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Orange"#"Red"

ds.camp[ds.camp$data >=
                          as.POSIXct("2020-12-28 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS") & ds.camp$data <=
                          as.POSIXct("2020-12-29 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Orange"#"Yellow"

ds.camp[ds.camp$data >=
                          as.POSIXct("2020-12-30 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS") & ds.camp$data <=
                          as.POSIXct("2021-01-03 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Orange"#"Red"

ds.camp[ds.camp$data ==
                          as.POSIXct("2021-01-04 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Orange"#"Yellow"

ds.camp[ds.camp$data >=
                          as.POSIXct("2021-01-05 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS") & ds.camp$data <=
                          as.POSIXct("2021-01-06 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Orange"#"Red"

ds.camp[ds.camp$data >
                          as.POSIXct("2021-01-06 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Yellow"

ds.camp[ds.camp$data >=
                          as.POSIXct("2021-01-09 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Yellow"#"Orange"

ds.camp[ds.camp$data >=
                          as.POSIXct("2021-01-11 00:00:00", 
                          format = "%Y-%m-%d %H:%M:%OS"),]$color <- "Yellow"





ds.camp$color=as.factor(ds.camp$color)
ds.camp$tamp_rate=ds.camp$nuovi_positivi/ds.camp$casi_testati
ds.camp$tamp_rate_variazione=ds.camp$nuovi_positivi_variazione/ds.camp$casi_testati_variazione



```{r fig.height=13,fig.width=13}
library(DataCombine)

date <- function(dd=1, mm, yy=2020){
  text = paste(yy, "-", mm, "-", dd, " 00:00:00", sep="")
  as.POSIXct(text, format = "%Y-%m-%d %H:%M:%OS")
}



d=ds.camp
d_test = ds.camp
#d_test$color = shifted(d_test$color, 14)
#d_test$rt = shifted(d_test$rt, 14)
d_test$temp = shifted(d_test$ricoverati_con_sintomi, -14)
d_test = d_test[d_test$data <= date(15,2,2021) & d_test$data >= date(1,9,2020), ]

#d$color=shifted(d$color,14)
#d$rt=shifted(d$rt, 14)
d$temp=shifted(d$ricoverati_con_sintomi, -14)
d=d[d$data<=date(26,01,2021) & d$data>=date(1,9,2020), ]

d$rt2 <- 1/d$rt
d_test$rt2 <- 1/d_test$rt
```

```{r, echo = TRUE}

d$temp=shifted(d$nuovi_positivi, -14)
d_test$temp=shifted(d_test$nuovi_positivi, -14)


m_nuovi_positivi_col = glm(data=d, formula = temp ~ 
                             tamp_rate:color +
                             I(tamp_rate^2) +  
                             terapia_intensiva +
                             I(isolamento_domiciliare) +
                             I(rt^2) +
                             rt:color,
                             #splines::ns(driving, 3),
                           family = Gamma(link="log")
                           )

#1596
summary(m_nuovi_positivi_col)

plot(d$temp)
lines(predict.glm(m_nuovi_positivi_col, d_test, type = "response"), col="blue")
lines(m_nuovi_positivi_col$fitted.values, col="red")

m_nuovi_positivi_col = glm(data=d, formula = temp ~ 
                             tamp_rate:color +
                             I(tamp_rate^2) +  
                             I(isolamento_domiciliare) +
                             I(rt^2) +
                             rt:color +
                             splines::ns(driving, 3) +
                             driving:isolamento_domiciliare +
                             driving:terapia_intensiva,
                           family = Gamma(link="log")
                           )

#1556
summary(m_nuovi_positivi_col)

plot(d$temp)
lines(predict.glm(m_nuovi_positivi_col, d_test, type = "response"), col="blue")
lines(m_nuovi_positivi_col$fitted.values, col="red")

m_nuovi_positivi_col = glm(data=d, formula = temp ~ 
                             tamp_rate:color +
                             I(tamp_rate^2) +  
                             I(isolamento_domiciliare) +
                             I(rt^2) +
                             rt:color +
                             splines::ns(driving, 3) +
                             driving:isolamento_domiciliare +
                             driving:terapia_intensiva +
                             driving:color,
                           family = Gamma(link="log")
                           )

#1551
summary(m_nuovi_positivi_col)

plot(d$temp)
lines(predict.glm(m_nuovi_positivi_col, d_test, type = "response"), col="blue")
lines(m_nuovi_positivi_col$fitted.values, col="red")


m_nuovi_positivi_col = glm(data=d, formula = temp ~ 
                             tamp_rate:color +
                             I(tamp_rate^2) +  
                             I(isolamento_domiciliare) +
                             I(rt^2) +
                             rt:color +
                             splines::ns(driving, 5) +
                             driving:isolamento_domiciliare +
                             driving:terapia_intensiva,
                           family = Gamma(link="log")
                           )

#1554
summary(m_nuovi_positivi_col)

plot(d$temp)
lines(predict.glm(m_nuovi_positivi_col, d_test, type = "response"), col="blue")
lines(m_nuovi_positivi_col$fitted.values, col="red")

```




```{r glm, echo = TRUE}
d$temp=shifted(d$ricoverati_con_sintomi, -14)
d_test$temp=shifted(d_test$ricoverati_con_sintomi, -14)


m_ricoverati=glm(data=d,formula=temp~
                  ricoverati_con_sintomi:color +
                  ricoverati_con_sintomi_variazione +
                  rt, 
                  family = Gamma(link="log"))
summary(m_ricoverati)
d_test$temp2 = shifted(d_test$ricoverati_con_sintomi, -14)
plot(d_test$temp2)
lines(predict.glm(m_ricoverati, d_test, type = "response"), col="blue")
lines(m_ricoverati$fitted.values, col="red")
#1414.2


#rt^(-1)
d$rt2 <- 1/d$rt
m_ricoverati=glm(data=d,formula=temp~
                  ricoverati_con_sintomi:color +
                  ricoverati_con_sintomi_variazione + 
                  nuovi_positivi + 
                  rt2, 
                  family = Gamma(link="log"))
summary(m_ricoverati)
d_test$temp2 = shifted(d_test$ricoverati_con_sintomi, -14)
plot(d_test$temp2)
lines(predict.glm(m_ricoverati, d_test, type = "response"), col="blue")
lines(m_ricoverati$fitted.values, col="red")
#1463.4

m_ricoverati=glm(data=d,formula=temp~
                  ricoverati_con_sintomi:color +
                  ricoverati_con_sintomi_variazione + 
                  nuovi_positivi + 
                  rt, 
                  family = Gamma(link="log"))
summary(m_ricoverati)
d_test$temp2 = shifted(d_test$ricoverati_con_sintomi, -14)
plot(d_test$temp2)
lines(predict.glm(m_ricoverati, d_test, type = "response"), col="blue")
lines(m_ricoverati$fitted.values, col="red")
#1413.5

d$temp=shifted(d$ricoverati_con_sintomi, -14)
m_ricoverati=glm(data=d,formula=temp~
                  poly(terapia_intensiva, 3, raw=TRUE) +
                  poly(isolamento_domiciliare, 3, raw=TRUE) +
                  rt:ricoverati_con_sintomi +
                  rt:color, 
                  family = Gamma(link="log"))
summary(m_ricoverati)
d_test$temp2 = shifted(d_test$ricoverati_con_sintomi, -14)
plot(d_test$temp2)
lines(predict.glm(m_ricoverati, d_test, type = "response"), col="blue")
lines(m_ricoverati$fitted.values, col="red")
#1264

d$temp=shifted(d$ricoverati_con_sintomi, -14)
m_ricoverati=glm(data=d,formula=temp~
                  splines::ns(terapia_intensiva, 4) +
                  splines::ns(isolamento_domiciliare, 4) +
                  splines::ns(ricoverati_con_sintomi, 4) +
                  rt:color +
                  color, 
                  family = Gamma(link="log"))
summary(m_ricoverati)
d_test$temp2 = shifted(d_test$ricoverati_con_sintomi, -14)
plot(d_test$temp2)
lines(predict.glm(m_ricoverati, d_test, type = "response"), col="blue")
lines(m_ricoverati$fitted.values, col="red")
#1159

d$temp=shifted(d$ricoverati_con_sintomi, -14)
m_ricoverati=glm(data=d,formula=temp~
                  splines::ns(terapia_intensiva, 4) +
                  splines::ns(isolamento_domiciliare, 4) +
                  splines::ns(ricoverati_con_sintomi, 4) +
                  rt:color +
                  color +
                  #splines::ns(driving, 2) +
                  rt:driving, 
                  family = Gamma(link="log"))
summary(m_ricoverati)
d_test$temp2 = shifted(d_test$ricoverati_con_sintomi, -14)
plot(d_test$temp2)
lines(predict.glm(m_ricoverati, d_test, type = "response"), col="blue")
lines(m_ricoverati$fitted.values, col="red")
#1227 ma miglior fit

#Aumento significatività

d$temp=shifted(d$ricoverati_con_sintomi, -14)
m_ricoverati=glm(data=d,formula=temp~
                  splines::ns(isolamento_domiciliare, 3) +
                  splines::ns(terapia_intensiva, 4) +
                  rt:color +
                  #color +
                  splines::ns(driving, 3) +
                  rt:driving, 
                  family = Gamma(link="log"))
summary(m_ricoverati)
d_test$temp2 = shifted(d_test$ricoverati_con_sintomi, -14)
plot(d_test$temp2)
lines(predict.glm(m_ricoverati, d_test, type = "response"), col="blue")
lines(m_ricoverati$fitted.values, col="red")

```






```{r, echo = TRUE}

d$temp=shifted(d$terapia_intensiva, -14)
m_terapia_intensiva=glm(data=d,formula=temp~
                  ricoverati_con_sintomi +
                  splines::ns(isolamento_domiciliare, 5) +
                  splines::ns(terapia_intensiva, 5), 
                  family = Gamma(link="log"))
summary(m_terapia_intensiva)
d_test$temp2 = shifted(d_test$terapia_intensiva, -14)
plot(d_test$temp2)
lines(predict.glm(m_terapia_intensiva, d_test, type = "response"), col="blue")
lines(m_terapia_intensiva$fitted.values, col="red")
#736


d$temp=shifted(d$terapia_intensiva, -14)
m_terapia_intensiva=glm(data=d,formula=temp~
                  splines::ns(isolamento_domiciliare, 5) +
                  splines::ns(terapia_intensiva, 5) +
                  splines::ns(tamp_rate, 5) +
                  splines::ns(nuovi_positivi, 3), 
                  family = Gamma(link="log"))
summary(m_terapia_intensiva)
d_test$temp2 = shifted(d_test$terapia_intensiva, -14)
plot(d_test$temp2)
lines(predict.glm(m_terapia_intensiva, d_test, type = "response"), col="blue")
lines(m_terapia_intensiva$fitted.values, col="red")
#708

d$temp=shifted(d$terapia_intensiva, -14)
m_terapia_intensiva=glm(data=d,formula=temp~
                  splines::ns(isolamento_domiciliare, 5) +
                  splines::ns(terapia_intensiva, 5) +
                  splines::ns(tamp_rate, 5) +
                  splines::ns(nuovi_positivi, 3) +
                  splines::ns(driving, 2), 
                  family = Gamma(link="log"))
summary(m_terapia_intensiva)
d_test$temp2 = shifted(d_test$terapia_intensiva, -14)
plot(d_test$temp2)
lines(predict.glm(m_terapia_intensiva, d_test, type = "response"), col="blue")
lines(m_terapia_intensiva$fitted.values, col="red")
#676

#Aumento variabili con significatività

d$temp=shifted(d$terapia_intensiva, -14)
m_terapia_intensiva=glm(data=d,formula=temp~
                  splines::ns(isolamento_domiciliare, 3) +
                  splines::ns(terapia_intensiva, 3) +
                  #splines::ns(tamp_rate, 2) +
                  splines::ns(nuovi_positivi, 3) +
                  splines::ns(driving, 3), 
                  family = Gamma(link="log"))
summary(m_terapia_intensiva)
d_test$temp2 = shifted(d_test$terapia_intensiva, -14)
plot(d_test$temp2)
lines(predict.glm(m_terapia_intensiva, d_test, type = "response"), col="blue")
lines(m_terapia_intensiva$fitted.values, col="red")
#797

d$temp=shifted(d$terapia_intensiva, -14)
m_terapia_intensiva=glm(data=d,formula=temp~
                  splines::ns(isolamento_domiciliare, 3) +
                  splines::ns(terapia_intensiva, 3) +
                  splines::ns(nuovi_positivi, 3) +
                  splines::ns(driving, 3) +
                  driving:tamp_rate +
                  driving:rt, 
                  family = Gamma(link="log"))
summary(m_terapia_intensiva)
d_test$temp2 = shifted(d_test$terapia_intensiva, -14)
plot(d_test$temp2)
lines(predict.glm(m_terapia_intensiva, d_test, type = "response"), col="blue")
lines(m_terapia_intensiva$fitted.values, col="red")
#784

d$temp=shifted(d$terapia_intensiva, -14)
m_terapia_intensiva=glm(data=d,formula=temp~
                  splines::ns(isolamento_domiciliare, 3) +
                  splines::ns(terapia_intensiva, 3) +
                  splines::ns(nuovi_positivi, 3) +
                  splines::ns(driving, 3) +
                  driving:nuovi_positivi, 
                  family = Gamma(link="log"))
summary(m_terapia_intensiva)
d_test$temp2 = shifted(d_test$terapia_intensiva, -14)
plot(d_test$temp2)
lines(predict.glm(m_terapia_intensiva, d_test, type = "response"), col="blue")
lines(m_terapia_intensiva$fitted.values, col="red")
#792

#NOTA: TUTTI GLI AIC DEVONO ESSERE CONSIDERATI SE IL MODELLO FITTA FINO AL PRIMO GENNAIO
```



`


```{r}
d$temp=shifted(d$isolamento_domiciliare, -14)
m_isolamento_domiciliare=glm(data=d,formula=temp~
                  splines::ns(isolamento_domiciliare, 5) +
                  splines::ns(terapia_intensiva, 5) +
                  splines::ns(tamp_rate, 5) +
                  splines::ns(nuovi_positivi, 3) +
                  rt:color +
                  color, 
                  family = Gamma(link="log"))
summary(m_isolamento_domiciliare)
d_test$temp2 = shifted(d_test$isolamento_domiciliare, -14)
plot(d_test$temp2)
lines(predict.glm(m_isolamento_domiciliare, d_test, type = "response"), col="blue")
lines(m_isolamento_domiciliare$fitted.values, col="red")

d$temp=shifted(d$isolamento_domiciliare, -14)
m_isolamento_domiciliare=glm(data=d,formula=temp~
                  splines::ns(isolamento_domiciliare, 5) +
                  splines::ns(terapia_intensiva, 5) +
                  splines::ns(tamp_rate, 5) +
                  splines::ns(nuovi_positivi, 3) +
                  rt:color +
                  color, 
                  family = Gamma(link="log"))
summary(m_isolamento_domiciliare)
d_test$temp2 = shifted(d_test$isolamento_domiciliare, -14)
plot(d_test$temp2)
lines(predict.glm(m_isolamento_domiciliare, d_test, type = "response"), col="blue")
lines(m_isolamento_domiciliare$fitted.values, col="red")
#2102

d$temp=shifted(d$isolamento_domiciliare, -14)
m_isolamento_domiciliare=glm(data=d,formula=temp~
                  splines::ns(isolamento_domiciliare, 5) +
                  splines::ns(terapia_intensiva, 5) +
                  splines::ns(tamp_rate, 5) +
                  splines::ns(nuovi_positivi, 3) +
                  rt:color +
                  color +
                  rt:driving +
                  nuovi_positivi:driving, 
                  family = Gamma(link="log"))
summary(m_isolamento_domiciliare)
d_test$temp2 = shifted(d_test$isolamento_domiciliare, -14)
plot(d_test$temp2)
lines(predict.glm(m_isolamento_domiciliare, d_test, type = "response"), col="blue")
lines(m_isolamento_domiciliare$fitted.values, col="red")
#2077

d$temp=shifted(d$isolamento_domiciliare, -14)
m_isolamento_domiciliare=glm(data=d,formula=temp~
                  #splines::ns(isolamento_domiciliare, 5),
                  splines::ns(terapia_intensiva, 5) + #tenere
                  splines::ns(isolamento_domiciliare, 5) +
                  #splines::ns(tamp_rate, 5) +
                  splines::ns(isolamento_domiciliare,4),
                  #rt:color +
                  #rt:driving +
                  #nuovi_positivi:driving, 
                  family = Gamma(link="log"))
summary(m_isolamento_domiciliare)
d_test$temp2 = shifted(d_test$isolamento_domiciliare, -14)
plot(d_test$temp2)
lines(predict.glm(m_isolamento_domiciliare, d_test, type = "response"), col="blue")
lines(m_isolamento_domiciliare$fitted.values, col="red")
#2077
```


## Explanatory analysis


We want to study `totale_positivi` in `Campania`. We should first plot some time-related data. `totale_positivi` represents the daily total of people that tested positive to COVID-19 and weren't already positive. PCRs are also used on already positive people that want to test if they still are. These last type positives doesn't add to the count. 

Color: 

* 6 Nov to 14 Nov  - Orange 
* 15 Nov to 5 Dec - Red 
* 6 Dec to 23 Dec - Orange 
* 24 dec to 27 dec - Red 
* 28 to 29 - Yellow 
* 30 to 3 - Red
* 4 - Yellow 
* 5 to 6 Jan - Red 
* 7 to ... - Yellow 

Important events:

* 3 Ott - mandatory masks and 22:00 curfew
* 15 Ott - De Luca closed schools (but nidi and asili)
* 24 Ott - protests against curfew
* 30 Ott - De Luca closed nidi and asili
* 24 Nov - gradual reopening of schools
* 25 Nov - Maradona memorial with thousands of people
* 25 Dec - Christmas
* 31 Dec - New Year's eve

## Quality of data

Data is collected daily. The first consideration about the quality of data is related to the periodicity of drops in the number of daily cases and number of daily deaths. The PCR tests usually give a response in 48h. During the weekend less tests are carried and this results in less cases during mondays and tuesdays. The other weekdays are also scattered and the maximum is usually reached on a wednesday. To minimize this effect, a 7-day average could be used.

A possible model for this phenomenon with the variables that we have at disposal could make use of `data`.

Possible relationships between `data` and `totale_positivi` could be polynomial or, given the exponential nature of the phenomenon, even logistic.



```{r}
selection <- c("data","nuovi_positivi", "isolamento_domiciliare", "totale_ospedalizzati", "positivi_ieri", "dimessi_guariti", "deceduti", "tamp_rate", "tamp_rate_ieri")
d2 <- d[selection]



for (i in c("isolamento_domiciliare", "totale_ospedalizzati", "positivi_ieri", "dimessi_guariti", "deceduti")){
  plot(d2[i], d2$nuovi_positivi, 
       col=time_interval(d2$data),
       main=i)
}
```


```{r}
plot(d2$tamp_rate_ieri, d2$tamp_rate - d2$tamp_rate_ieri, col=time_interval(d2$data, interval=10))
abline(h=0)
#plot(d2$nuovi_positivi, d2$positivi_ieri, col=d2$color)
#lines(d2$tamp_rate_ieri, 0)
```

```{r}
plot(d2$data, d2$tamp_rate_ieri, col=time_interval(d2$data, interval=10))
abline(h=0)
```

```{r fig.width=9,fig.height=9}

model_positivi=glm(data=d2,formula=nuovi_positivi~isolamento_domiciliare+totale_ospedalizzati+positivi_ieri+tamp_rate_ieri,  family = Gamma(link="log"))
summary(model_positivi)
#par(mfrow=c(2,2))
plot(model_positivi)
plot(model_positivi$fitted.values)
points(d2$nuovi_positivi,col="red")


```






